### 2. Visualization - work on your own data
- install necessary packages for this session

```{r, message=FALSE}
BiocManager::install('devtools', ask = F, update = F)
devtools::install_github("EdwinTh/dutchmasters")
BiocManager::install('ggsci',ask = F, update = F)
```

- load necessary packages for this session

```{r, message=FALSE}
library(phyloseq)
library(ggplot2)
library(tidyverse)
library(ggpubr)
library(ggsci)
library(stringr)
library(dutchmasters)
```

##### 2.1 data structure of phyloseq object

```{r, message=FALSE}
load('phyloseq_object_data_2.Rdata')
```

```{r}
phylo_obj_2@sam_data[1:3,1:3]
```

```{r}
phylo_obj_2@otu_table[1:3,1:3]
```

```{r}
phylo_obj_2@tax_table[1:3,1:3]
```

##### 2.2 convert the three tables to a combined data frame for plotting using function psmelt()

```{r}
pd <- psmelt(phylo_obj_2)
pd[1:7,1:7]
```
- 1890 obs. of 13 variables

```{r}
pd = pd %>% mutate(Kingdom = str_replace(Kingdom, 'k__',''),
                   Phylum = str_replace(Phylum, 'p__',''),
                   Class = str_replace(Class, 'c__',''),
                   Order = str_replace(Order, 'o__',''),
                   Family = str_replace(Family, 'f__',''),
                   Genus = str_replace(Genus, 'g__',''),
                   Species = str_replace(Species, 's__',''))
```

#### 2.3 which columns and which function will be used for bar plot?

```{r}
colnames(pd)
```

#### # create a basic bar plot

```{r}
# The abundance of species in each sample
ggplot(data = pd, aes(x = SampleID, y = Abundance)) + 
  geom_bar(stat = 'identity')
```

#### Q:the label of sampleID was stacked, how to solve it?

```{r}
ggplot(data = pd, aes(x = SampleID, y = Abundance)) + 
  theme_bw() +
  geom_bar(stat = 'identity') + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

#### Q: how to add group information

```{r}
ggplot(data = pd, aes(x = SampleID, y = Abundance)) + 
  geom_bar(stat = 'identity', aes(fill = Condition)) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

#### Q: how to change width of bar

```{r}
ggplot(data = pd, aes(x = SampleID, y = Abundance)) + 
  geom_bar(stat = 'identity', aes(fill = Condition),width = 0.2) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

#### Q: how to custom the color of bar

```{r}
ggplot(data = pd, aes(x = SampleID, y = Abundance)) + 
  geom_bar(stat = 'identity', aes(fill = Condition),width = 0.8) + 
  #scale_fill_manual(values = c("blue","red","green")) +
  scale_fill_manual(values = c("#b3e2cd","#fdcdac","#cbd5e8")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

#### Q: The distribution of class in those groups

```{r}
ggplot(data = pd, aes(x = Class, y = Abundance)) + 
  geom_bar(stat = 'identity', aes(fill = Condition),width = 0.8) + 
  #scale_fill_manual(values = c("blue","red","green")) +
  scale_fill_manual(values = c("#b3e2cd","#fdcdac","#cbd5e8")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

#### Q: how to change the colors for mutiple classes

```{r}
#length(unique(pd$Phylum))
#6
ggplot(data = pd, aes(x = Condition, y = Abundance)) + 
  geom_bar(stat = 'identity', position = "fill", aes(fill = Phylum),width = 0.8) + 
  scale_fill_jco()
```


```{r}
#length(unique(pd$Class))
#13
ggplot(data = pd, aes(x = Condition, y = Abundance)) + 
  geom_bar(stat = 'identity', position = "fill", aes(fill = Class),width = 0.8) + 
  scale_fill_dutchmasters(palette = "milkmaid")
```

#### Q: how to show relative abundance of class in each group
- add position = "fill" with geom_bar()
```{r}
ggplot(data = pd, aes(x = Class, y = Abundance)) + 
  geom_bar(stat = 'identity', position = "fill", aes(fill = Condition),width = 0.8) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_jco()
```

#### Q: how to change the title of y-axis

```{r}
ggplot(data = pd, aes(x = Class, y = Abundance)) + 
  geom_bar(stat = 'identity', position = "fill", aes(fill = Condition),width = 0.8) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_jco() +
  labs(x = "Group", y = "Relative abundance (%)")
```

#### Q: how to change the backgroud

```{r}
ggplot(data = pd, aes(x = Class, y = Abundance)) + 
  geom_bar(stat = 'identity', position = "fill", aes(fill = Condition),width = 0.8) + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_jco() +
  labs(x = "Group", y = "Relative abundance (%)")
```

#### Q: how to change the position of legend.

```{r}
ggplot(data = pd, aes(x = Condition, y = Abundance)) + 
  geom_bar(stat = 'identity', position = "fill", aes(fill =Phylum ),width = 0.8) + 
  theme_classic() +
  theme(legend.position = 'top') +
  scale_fill_jco() +
  labs(x = "Group", y = "Relative abundance (%)")
```

#### Q: how to save your figure
- use ggsave()
```{r}
fig <- ggplot(data = pd, aes(x = Condition, y = Abundance)) + 
  geom_bar(stat = 'identity', position = "stack", aes(fill =Phylum ),width = 0.8) + 
  theme_classic() +
  theme(legend.position = 'top') +
  scale_fill_jco() +
  labs(x = "Group", y = "Relative abundance (%)")
ggsave(fig, filename = 'Con2Abundance.pdf', width = 5, height = 4)
```

#### 2.4 which columns and which function will be used for box plot?
- use geom_boxplot()
```{r}
load('Data2_Abundance_metadata.Rdata')
head(Metadata)
```


```{r}
ggplot(data = Metadata, aes(x = Condition, y = Shannon)) + 
  geom_boxplot(aes(fill = Condition))
```

#### Q: how to evalute the difference of Shannon between groups, is it significant?

```{r}
my_comparisons = list(c("Blautia-HFD","Control-HFD"),c("Control-HFD","Control-Chow",""))
ggplot(data = Metadata, aes(x = Condition, y = Shannon)) + 
  geom_boxplot(aes(fill = Condition)) +
  scale_fill_jco() +
  stat_compare_means(comparisons = my_comparisons, method = "wilcox.test",aes(as_label(..p.format..))) # ..p.signif..
```


#### 2.5 heatmap using phyloseq package

```{r}
plot_heatmap(phylo_obj_2, method = "NMDS", distance = "bray")
```

#### Q: how to show the most abundant OTUs

```{r}
total = median(sample_sums(phylo_obj_2))
abundant_OTU <- filter_taxa(phylo_obj_2, function(x) sum(x > total*0.20) > 0, TRUE)
abundant_OTU
```

```{r}
pd_abundant  = psmelt(abundant_OTU)
head(pd_abundant)
```

```{r}
plot_heatmap(abundant_OTU, method = "NMDS", distance = "bray")
```


```{r}
plot_heatmap(abundant_OTU, method = "MDS", distance = "(A+B-2*J)/(A+B-J)", 
               taxa.label = "Order", taxa.order = "Order", 
               trans=NULL, low="beige", high="red", na.value="beige")
```


#### 2.6 heatmap in ggplot2

```{r}
ggplot(data = pd_abundant, aes(x=SampleID, y=Class, fill = Abundance)) +
  geom_tile() +
  scale_fill_continuous(low = 'beige', high = 'red') + 
  theme(axis.text.x = element_text(angle = 270, vjust = 0.5))
```

- it is more flexible using ggplot.

#### 2.7 Simple network analysis

```{r}
plot_net(phylo_obj_2, distance = "(A+B-2*J)/(A+B)", type = "taxa",
         maxdist = 0.7, color="Class", point_label="Genus")
```




